### 2023-08-19 code/analysis-pilot.Rmd
new labels

---
title: "analysis-pilot"
output: html_document
date: "2023-07-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

pilot_start=lubridate::ymd('2023-07-07') 
pilot_end=lubridate::ymd('2023-07-12')

real_start=lubridate::ymd('2023-07-17')
real_end=lubridate::ymd('2023-08-09')

data_source <- "data/rounds.csv"
player_source <- "data/players.csv"

pilot_ids <- str_c("id", 3:12) # Ids 1, 2 were adult pilot test

real_ids <-str_c("id", c(13:16, 19:36, 39:54, 57:58)) # skipping kids who didn't complete enough trials
```

# Data prep

```{r}

players <- read_csv(here(player_source)) |> filter(id %in% real_ids) %>% select(id, playerId=`_id`)

games <- read_csv(here("data/games.csv"), col_types="ccc")|> filter(is.na(exclude)) |>  mutate(offset_sec=lubridate::ms(offset) |> lubridate::period_to_seconds()) |> select(-offset)

```

```{r}

raw_data <- read.csv(here(data_source), header = TRUE) %>% 
  filter(createdAt >= real_start) %>%
  filter(createdAt <= real_end) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  select(-chat) |> 
  pivot_longer(cols=starts_with('player'), names_to=c("playerId", "info"), names_prefix="player_", names_sep="_",
               values_to="values", values_transform=as.character, values_drop_na=T) %>% 
  filter(values!="") %>% 
  filter(playerId!=speaker) %>% 
  pivot_wider(names_from=info, values_from=values) |> 
  left_join(players, by=c("playerId")) |> 
  select(-playerId) |> 
  rename(playerId=id) |> 
  left_join(players, by=c("speaker"="playerId")) |> 
  select(-speaker) |> 
  rename(speaker=id) %>% 
  left_join(games) |> 
  rename(game=gameConfig) |> 
  select(-gameId, -offset_sec) |> 
  filter(!is.na(game)) |> 
  mutate(time=as.numeric(time))

```

```{r}
#just for pilot, need to fix offset
#fix <- raw_data |> filter(trialNum==0) |> select(game, time)

#real_offset <- games |> left_join(fix, by=c("gameConfig"="game")) |> mutate(offset_sec=offset_sec-time/1000)

real_offset <- games

```

```{r}

wrangle_transcript <- function(game_num){
read_csv(here("data/clean_transcripts", str_c(game_num,".csv")), show_col_types=F) |> 
  select(start, end, text, speaker) |> 
  mutate(start=start/1000, end=end/1000) |> 
    nest(data=c(start, end, text, speaker))
}

#wrangle_transcript("game32")

real_offset$transcripts <- map_df(real_offset$gameConfig, wrangle_transcript)

transcripts <- real_offset |> unnest(transcripts) |> unnest(data)

# this section is janky as hell, but it does work

```

```{r}

stages <- read_csv(here("data/stages.csv")) |> select(index, gameId, startTimeAt) |> group_by(gameId)

gameStarts <- stages |> filter(index==0) |> rename(first=startTimeAt) |> select(-index)

stage_timing <- stages |> left_join(gameStarts) |>
  mutate(rel_time=startTimeAt-first) |> 
  mutate(rel_time=lubridate::as.period(rel_time) |> lubridate::period_to_seconds()) |> 
   mutate(trialNum=index/2,
          start=rel_time,
          end=lead(rel_time)) |> 
  select(gameId, trialNum,start, end) |>  left_join(games) |>  ungroup() |> 
  mutate(trialNumStart=trialNum, trialNumEnd=trialNum) |> 
  select(game=gameConfig, trialNumStart, trialNumEnd, start, end) |> filter(!is.na(game)) |> filter(!is.na(start)) |> filter(!is.na(end))
```

```{r}
timed_transcripts <-  transcripts |>
  mutate(start=start-offset_sec, end=end-offset_sec) |> 
  select(-offset_sec, -gameId) |> 
  mutate(trialNumStart=NA, trialNumEnd=NA) |> 
  rename(game=gameConfig) |> 
  bind_rows(stage_timing) |> 
  group_by(game) |> 
  arrange(start) |> 
  fill(trialNumStart) |> 
  arrange(end) |> 
  fill(trialNumEnd, .direction="up") |> 
  arrange(game) |> 
  filter(!is.na(text)) |> 
  write_csv(here("data/timed_transcript.csv"))
```

```{r}
# how many trial mismatches are there?
timed_transcripts |> filter(end>0) |> filter(!is.na(trialNumEnd)) |> #it's in the game region, not rules or post game
 mutate(mismatch=trialNumEnd-trialNumStart) |> 
  group_by(mismatch) |> 
  tally()

#so like a third mismatch, and only a couple off by more than 1
```
<!--
partner 1:
id13
id15
id19
id21
id24
id26
id28
id30
id32
id33
id35
id39
id42
id43
id45
id47
id50
id52
id54
id58

partner 2:
id14
id16
id20
id22
id23
id25
id27
id29
id31
id34
id36
id40
id41
id44
id46
id48
id49
id51
id53
id57
-->

# timed_transcripts editing for partner1
```{r}
# timed_new <- timed_transcripts %>%
#   filter(!speaker %in% c("I", "M", "i", "id14", "id16", "id20", "id22", "id23", "id25", "id27", "id29", "id31", "id34", "id36", "id40", "id41", "id44", "id46", "id48", "id49", "id51", "id53", "id57")) %>%
#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>%
#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>%
#   filter(!roundStart %in% c(2.0, 4.0, 6.0, 8.0, 10.0, 12.0)) %>%
#   group_by(game, roundStart) %>%
#     summarize(totwordCount = sum(wordCount))
# 
# View(timed_new)

```

# timed_new graphing - left in speaker1listener
```{r}
# ggplot(timed_new, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
#   geom_line() +
#   #geom_smooth() +
#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
#   scale_x_continuous(breaks=seq(3,13,2), limits = c(3,13)) +
#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
#   theme_classic() +
#   theme(aspect.ratio = 9/9,
#         axis.title=element_text(size=14,face="bold"),
#         axis.text=element_text(size=9, color = "black"),
#         legend.title=element_text(size=13,face="bold"),
#         legend.text=element_text(size=9),
#         legend.position="right")
```
# timed_transcripts editing for partner2
```{r}
# timed_neww <- timed_transcripts %>% 
#   filter(!speaker %in% c("I", "M", "i", "id13", "id15", "id19", "id21", "id24", "id26", "id28", "id30", "id32", "id33", "id35", "id39", "id42", "id43", "id45", "id47", "id50", "id52", "id54", "id58")) %>% 
#   filter(!trialNumStart %in% c(0.0, 0.5, 1.0, 13.0, NA)) %>% 
#   mutate(wordCount = str_count(text, "\\S+"), roundStart = ceiling(trialNumStart)) %>% 
#   filter(!roundStart %in% c(3.0, 5.0, 7.0, 9.0, 11.0, 13.0)) %>% 
#   group_by(game, roundStart) %>% 
#     summarize(totwordCount = sum(wordCount))
# 
# View(timed_neww)

```

# timed_neww graphing - left in speaker2listener
```{r}
# ggplot(timed_neww, aes(x = roundStart, y = totwordCount, group = game, color = game)) + 
#   geom_line() +
#   #geom_smooth() +
#   labs(color = "Game ID", x = "\nRound #", y = "Utterance Count\n") +
#   scale_x_continuous(breaks=seq(2,12,2), limits = c(2,12)) +
#   scale_y_continuous(breaks=seq(0,30,1), limits = c(0,30)) +
#   theme_classic() +
#   theme(aspect.ratio = 9/9,
#         axis.title=element_text(size=14,face="bold"),
#         axis.text=element_text(size=9, color = "black"),
#         legend.title=element_text(size=13,face="bold"),
#         legend.text=element_text(size=9),
#         legend.position="right")
```

# Prep accuracy

## remove columns: numPlayers, sound, activePlayerCount, trialNum, distractor, tangramURLs, response, target, submitted, X_id, repNum, correct, targetNum, stageIds
```{r}
filtered_data <- raw_data %>% 
  select(-numPlayers) %>% 
  select(-sound) %>%  
  select(-activePlayerCount) %>% 
  select(-trialNum) %>% 
  select(-distractor) %>% 
  select(-tangramURLs) %>% 
  select(-response) %>% 
  select(-submitted) %>% 
  select(-X_id) %>% 
  select(-repNum) %>% 
  select(-correct) %>% 
  select(-targetNum) %>% 
  select(-stageIds)

```

## remove NA cells in "time"
```{r}
filtered_data <- filtered_data %>% 
  drop_na(time)

View(filtered_data)
```

## relocate columns
```{r}
filtered_data <- filtered_data %>% 
  select(createdAt, everything()) %>% 
  relocate(countCorrect, .before = time)

#View(filtered_data)
```

# accuracy graph
```{r}
accuracy <- filtered_data %>% 
  filter(!index %in% c(0, 1)) %>%
  group_by(game) %>% 
  summarize(total=n(),
            correct=sum(countCorrect)) %>% 
  mutate(fraction = correct/total*100, percent = signif(fraction, 2))

ggplot(data = accuracy, aes(x = percent, y = reorder(game, -percent))) +
  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nPercentage accuracy") +
  ylab("Game ID\n") +
  scale_x_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
ggsave("AccuracySorted.png")

```
#accuracy through time
```{r}
accuracy <- filtered_data %>% 
  filter(!index %in% c(0, 1)) %>%
  group_by(index) %>% 
  summarize(total=n(),
            correct=sum(countCorrect)) %>% 
  mutate(fraction = correct/total*100, percent = signif(fraction, 2))

ggplot(data = accuracy, aes(x = index-1, y = percent)) +
  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
ggsave("AccuracybyTrial.png")
```
#accuracy by target
```{r}
accuracy <- filtered_data %>% 
  mutate(image = case_when(
  target == "/experiment/B1.jpg" ~ "Walking",
  target == "/experiment/D1.jpg" ~ "Lying",
  target == "/experiment/E1.jpg" ~ "Jumping",
  target == "/experiment/I1.jpg" ~ "Holding",
)) %>% 
  filter(!index %in% c(0, 1)) %>%
  group_by(image) %>% 
  summarize(total=n(),
            correct=sum(countCorrect)) %>% 
  mutate(fraction = correct/total*100, percent = signif(fraction, 2))

ggplot(data = accuracy, aes(x = image, y = percent)) +
  geom_bar(stat = "identity", position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(method="lm")+
  theme_classic() +
  xlab("\nTarget") +
  ylab("Percent accuracy\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(breaks=seq(0,100,10), limits = c(0, 100), expand=c(0,0)) +
  geom_text(aes(label = percent), vjust = 1.7, color = "white", size = 3) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
ggsave("AccuracybyTarget.png")
```

#reduction
```{r}
join_data <- filtered_data %>% 
  mutate(image = case_when(
  target == "/experiment/B1.jpg" ~ "Walking",
  target == "/experiment/D1.jpg" ~ "Lying",
  target == "/experiment/E1.jpg" ~ "Jumping",
  target == "/experiment/I1.jpg" ~ "Holding",
)) %>% 
  rename(listener = playerId, trialNum = index) %>% 
  select(-createdAt, -exclude, -time) %>% 
  pivot_longer(cols = "listener":"speaker", names_to = "role", values_to = "speaker")
```

```{r}
timed_new <- timed_transcripts %>% 
  drop_na(trialNumStart, trialNumEnd) %>% 
  filter(!speaker %in% c("I", "M", "i")) %>% 
  mutate(wordCount = str_count(text, "\\S+"), trialNum = ceiling(trialNumStart)) %>% 
  full_join(
    join_data,
    by = c("speaker", "game", "trialNum")
  ) %>%
  mutate(wordCount=ifelse(is.na(wordCount), 0, wordCount)) %>% 
  select(speaker, game, trialNum, role, countCorrect, wordCount, image) %>% 
  filter(!trialNum %in% c(0.0, 1.0)) %>% 
   group_by(game, trialNum, role, image) %>% 
    summarize(totwordCount = sum(wordCount))
```

#speaker graphing
```{r}
timed_speaker <- timed_new %>% 
  filter(role == "speaker")

ggplot(timed_speaker, aes(x = trialNum-1, y = totwordCount)) + 
  geom_line(aes(group=game)) +
  geom_smooth() +
  labs(color = "Game ID", x = "\nTrial", y = "Word Count\n") +
  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
  theme_classic() +
  theme(aspect.ratio = 9/9,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.title=element_text(size=13,face="bold"),
        legend.text=element_text(size=9),
        legend.position="right")
ggsave("SpeakerWord.png")
```

#listener graphing
```{r}
timed_listener <- timed_new %>% 
  filter(role == "listener")

ggplot(timed_listener, aes(x = trialNum-1, y = totwordCount)) + 
  geom_line(aes(group=game)) +
  geom_smooth() +
  labs(color = "Game ID", x = "\nTrial", y = "Word Count\n") +
  scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
  theme_classic() +
  theme(aspect.ratio = 9/9,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.title=element_text(size=13,face="bold"),
        legend.text=element_text(size=9),
        legend.position="right")
ggsave("ListenerWord.png")
```

<!-- #speakerbytarget -->
```{r}
# speaker_target <- timed_new %>% 
#   mutate(block=(trialNum-2)%/%4+1) %>% 
#   filter(role == "speaker") %>% 
#   mutate(gameximage=str_c(game,image))
# 
# ggplot(speaker_target, aes(x = block, y = totwordCount, color=image)) + 
#   #geom_line(aes(group=gameximage), alpha=.5, color="gray") +
#   #geom_jitter()+
#   geom_smooth(method="lm") +
#   labs(color = "Game ID", x = "\nRound #", y = "Word Count\n") +
#   #scale_x_continuous(breaks=seq(1,12,1), limits = c(1,12)) +
#   theme_classic() +
#   #facet_wrap(~image)+
#   theme(aspect.ratio = 9/9,
#         axis.title=element_text(size=14,face="bold"),
#         axis.text=element_text(size=9, color = "black"),
#         legend.title=element_text(size=13,face="bold"),
#         legend.text=element_text(size=9),
#         legend.position="right")

```