### 2023-09-29 code/analysis_1.Rmd
start data prep and analyses

---
title: "Analysis - expt 1"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)

transcript <- read_csv(here("data/timed_transcript.csv"))

dat <- read_csv(here("data/clean_data.csv"))

time_palette <- c("practice"="grey","block 1"="blue","block 2"="green", "block 3"="red")#TODO FIX
```

Trials were excluded if the speaker did not provide any description before the listener clicked.
# Accuracy over time

```{r}
# want to know which trials to exclude b/c no description given
exclusion <- transcript %>% filter(!is.na(description)) %>% filter(role=="S") %>% select(game, trialNum=trial, echo) %>% unique() %>% mutate("S_talked"=1)

dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  ggplot(aes(x = trialNum+1, y = as.numeric(correct), fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```
# Rate of echoing

will need to check more closely, may have missed some
```{r}
dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  mutate(is.echo=ifelse(is.na(echo),0,1)) %>% 
  ggplot(aes(x = trialNum+1, y = is.echo, fill=type)) +
  stat_summary(fun.data="mean_cl_boot", geom="bar") +
    stat_summary(fun.data="mean_cl_boot", geom="linerange", color="black") +
  scale_fill_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")
```

# Speed to start of description

```{r}

transcript %>% filter(role=="S") %>% filter(!is.na(description)) %>% group_by(game, trial) %>% 
  mutate(diff=end_time-trial_start) %>% summarize(first=min(diff)) %>% 
  mutate(type=case_when(trial<2 ~ "practice",
                        trial<6 ~ "block 1",
                        trial<10 ~ "block 2",
                        trial<14 ~ "block 3")) %>% 
  ggplot(aes(x = trial+1, y = first, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Time to start of description") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  #scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

TO DO try to understand negative outliers?

# Speed to response

```{r}

 dat %>% left_join(exclusion) %>% filter(S_talked==1) %>% 
  mutate(type=case_when(trialNum<2 ~ "practice",
                        trialNum<6 ~ "block 1",
                        trialNum<10 ~ "block 2",
                        trialNum<14 ~ "block 3")) %>% 
  ggplot(aes(x = trialNum+1, y = time/1000, color=type)) +
  geom_jitter(width=.2, height=0, alpha=.5)+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black") +
  scale_color_manual(values=time_palette)+
  #geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  #geom_smooth(method="lm")+
  #geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  coord_cartesian(ylim=c(0,100))+
  #ylab("Percent trials with echoing \n by experimenter\n") +
  #scale_x_continuous(breaks=seq(1,12,1)) +
  scale_y_continuous(breaks=seq(0,100,10), expand=c(0,0)) +
  theme(aspect.ratio = 4/5,
        axis.title=element_text(size=14,face="bold"),
        axis.text=element_text(size=9, color = "black"),
        legend.position = "none")

```

# Total kid words said in at least vague reference

# Descriptive something something
- qualitative on descriptions and similarities of descriptions (within/between kids) (maybe cut when kids say the same thing thrice down to once?)

- what descriptions "worked" -- some sort of qualitative range