### 2024-08-08 code/expt_1/leung_comparison.Rmd

---
title: "Leung data comparison"
output:
  html_document:
    df_print: paged
---
```{r prep, echo=F, message=F, warning=F}
library(here)
library(tidyverse)

knitr::opts_chunk$set(
                      echo=F, warning=F, cache=T, 
                      message=F, sanitize = T)
theme_set(theme_bw())
```

```{r, eval=F}
all_4yos <- list.files(here("leung_data/4yos")) |> map_df(~read_csv(here("leung_data/4yos/",.))) |> bind_rows() |> 
  write_csv(here("leung_data/4yos_uncoded.csv"))

all_adults <- list.files(here("leung_data/adult")) |> map_df(~read_csv(here("leung_data/adult/",.))) |> bind_rows() |> 
  write_csv(here("leung_data/adults_uncoded.csv"))

```

** note **: in the adult data, for sub id 104, trial 9 (target D1), it had been that two lines had selections marked. I believe the first should have had selection NA b/c it has text as well and this has been fixed so we don't have duplicates.

```{r}
kids <- read_csv(here("leung_data/4yos_coded_filtered.csv")) |> 
  filter(target %in% c("B1", "D1", "E1", "I1"))

adults <- read_csv(here("leung_data/adults_coded_filtered.csv"))|> 
  filter(target %in% c("B1", "D1", "E1", "I1"))

kids_selection <- kids |> filter(!is.na(selection)) |> select(listener=selection, target, trial, subid, type)

adults_selection <- adults |> filter(!is.na(selection)) |> select(listener=person, target, trial, subid, type)

adults_tagged <- adults |> left_join(adults_selection) |> 
  mutate(role=ifelse(person==listener, "listener", "speaker")) |> 
  mutate(words=str_count(description, "\\S+"),
         words=ifelse(is.na(words), 0, words),
         all_words=str_count(utterance, "\\S+"),
         all_words=ifelse(is.na(all_words), 0, all_words),
         block=(trial-1)%/%10)

kids_tagged <- kids |> 
  left_join(kids_selection) |> 
  mutate(role=ifelse(person==listener, "listener", "speaker")) |> 
   mutate(words=str_count(description, "\\S+"),
         words=ifelse(is.na(words), 0, words),
         all_words=str_count(utterance, "\\S+"),
         all_words=ifelse(is.na(all_words), 0, all_words),
         block=(trial-1)%/%10)

```

```{r, fig.height=3, fig.width=6}

adults_summary <- adults_tagged  |> 
  group_by(subid, role, block, target) |> 
  summarize(words=sum(words),
            all_words=sum(all_words))

  # ggplot(adults_summary, aes(x=block, y=words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,20))+
  # theme_classic()

  # ggplot(adults_summary, aes(x=block, y=all_words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,40))+
  # theme_classic()

  kids_summary <- kids_tagged  |> 
    filter(!is.na(role)) |> 
  group_by(subid, person, role, block, target) |> 
  summarize(words=sum(words),
            all_words=sum(all_words))

  #   ggplot(kids_summary, aes(x=block, y=words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", color="black")+
  #   coord_cartesian(ylim=c(0,20))+facet_grid(~ person)
  # 
  # 
  # ggplot(kids_summary, aes(x=block, y=all_words))+
  #   geom_jitter(width=.2, height=0, alpha=.5, color="blue")+
  #   stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
  #   coord_cartesian(ylim=c(0,40))+facet_grid(~person)

  all <- adults_summary |> mutate(person="adult-to-adult") |> 
    bind_rows(kids_summary) |> 
    mutate(person=case_when(
      person=="child" ~ "child-to-parent",
      person=="parent" ~ "parent-to-child",
      T ~ person
    ))

  time_palette <- c("practice"="grey","1"="#E41A1C","2"="#377EB8", "3"="#4DAF4A", "4"="purple")#TODO FIX
      ggplot(all |> filter(role=="speaker"), aes(x=block+1, y=words, color=as.factor(block+1)))+
    geom_jitter(width=.2, height=0, alpha=.4)+
    stat_summary(fun.data="mean_cl_boot", color="black")+
                stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
          scale_color_manual(values=time_palette)+
        theme(legend.position="none")+
    coord_cartesian(ylim=c(0,20), expand=F)+facet_grid(~ person)+
        labs(title="Re-analysis of Leung et al 2023 data", x="Block", y="Description length")

ggsave(here("foo.png"))

  ggplot(all |> filter(role=="speaker"), aes(x=block+1, y=all_words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
        stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,40))+facet_grid(~person)+
    labs(title="Total Words from speaker", x="Block", y="Total words")

        ggplot(all |> filter(role=="listener"), aes(x=block+1, y=words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", color="black")+
                stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,20))+facet_grid(~ person)+
        labs(title="Referential words from listener", x="Block", y="Referential words")

  ggplot(all |> filter(role=="listener"), aes(x=block+1, y=all_words))+
    geom_jitter(width=.2, height=0, alpha=.2, color="blue")+
    stat_summary(fun.data="mean_cl_boot", geom="pointrange", color="black")+
        stat_summary(fun.data="mean_cl_boot", geom="line", color="black")+
    coord_cartesian(ylim=c(0,40))+facet_grid(~person)+
    labs(title="Total Words from listener", x="Block", y="Total words")

```

### SBERT

```{r eval=F}
all <- adults_tagged |> bind_rows(kids_tagged) |> filter(!is.na(description)) |> filter(role=="speaker") |> 
  select(subid, target, person, description, block) %>%
  group_by(subid, block, target, person) |> filter(!is.na(description)) %>% 
  summarize(words=str_c(description, sep=" ", collapse=", ")) %>% write_csv(here("leung_data/pre_sbert.csv"))
```

```{r eval=F}
library(reticulate)
np <- import("numpy")
mat = np$load(here('leung_data/post_sbert.npy'))
saveRDS(mat,here('leung_data/post_sbert.RData'))
```

```{r}
sbert <- read_csv(here("leung_data/pre_sbert.csv")) %>% bind_cols(read_rds(here("leung_data/post_sbert.RData")) %>% 
  as_tibble())

### helper funcs
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}

### funcs
  F_mat <- sbert%>% select(starts_with("V")) %>% as.matrix() #Features
  M_mat <- sbert %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())

similarities <- M_mat %>%
    mutate(combinedId=str_c(subid,block,person,target,sep="_")) %>%
    make_across_df(F_mat, 'cosine') |> 
    separate(dim1, into=c("game1","block1","person1", "target1"), convert=T, sep="_") %>%
    separate(dim2, into=c("game2","block2","person2", "target2"), convert=T, sep="_") %>%
    mutate(sim = ifelse(is.nan(sim), NA, sim)) %>% write_rds(here("leung_data/similarities.rds"))

```

```{r, fig.width=5, fig.height=3}
sims <- read_rds(here("leung_data/similarities.rds")) |> as_tibble()

sims %>%  filter(!is.na(sim)) %>% 
  mutate(same_target=ifelse(target1==target2, "Same target", "Different target"),  
  same_child=case_when(
    game1!=game2 ~ "Different\n games",
    person1!=person2 ~ "Different\n speakers",
    T ~"Same\n speaker"),
  roles=case_when(
    person1 %in%c("left", "right") & person2 %in% c("left", "right") ~ "adult-adult",
    person1 == "parent" & person2=="parent" ~ "parent-parent",
    person1 == "child" & person2=="child" ~ "child-child",
    person1 == "child" & person2=="parent" ~ "parent-child",
    person1 == "parent" & person2=="child" ~ "parent-child",
    T ~ "bad"
  ) )|> 
    filter(roles!="bad") |> # don't want mix between adult & parent-child
  filter(block1!=block2) |> # can only compare across blocks in some cases, so only cross ever
  ggplot(aes(x=str_c(same_child), y=sim, color=roles)) +
  facet_grid(.~same_target) +
  theme_bw()+
  labs(x="", y="Cosine similarity", color="")+
  scale_color_brewer(type="qual", palette="Dark2")+
  scale_y_continuous(limits=c(0,1), expand=c(0,0))+
      stat_summary(fun.data="mean_cl_boot", geom="pointrange", position=position_dodge(width=.2))+
    guides( color = guide_legend( override.aes = list(linetype=0, size=1)))+
  theme(strip.background =element_rect(fill="white"),
        legend.position="bottom",
        axis.title.x=element_blank())

ggsave(here("bar.png"))

```