### 2025-08-07 code/demog.qmd
add demog (age in mo)

---
title: "demog"
format: html
---

```{r}
library(tidyverse)
library(here)
raw_demog <- read_csv(here("demographics.csv"))
```

```{r}
expt_1 <- read_csv(here("data/expt_1/clean_data.csv")) |>
  select(game, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 1) |>
  mutate(sid = id)

id_sub <- tribble(
  ~id, ~sid,
  "apple123", "id74",
  "pear124", "id75",
  "apple127", "id76",
  "pear128", "id77",
  "80", "id78",
  "81", "id79",
  "apple111", "id90",
  "pear222", "id91",
  "apple333", "id92",
  "pear444", "id93",
  "apple255", "id94",
  "pear257", "id95"
)

expt_2 <- read_csv(here("data/expt_2/clean_data.csv")) |>
  select(game = gameId, speaker, listener) |>
  pivot_longer(c("speaker", "listener"), values_to = "id") |>
  select(-name) |>
  unique() |>
  mutate(expt = 2) |>
  left_join(id_sub) |>
  mutate(sid = case_when(
    !is.na(sid) ~ sid,
    str_detect(id, "id") ~ id,
    T ~ str_c("id", id)
  ))

kids <- expt_1 |> bind_rows(expt_2) # included kids!
```

```{r}
demogs <- raw_demog |>
  mutate(
    date_test = mdy(date_test),
    date_birth = mdy(date_birth)
  ) |>
  select(sid, included, date_test, date_birth, hispanic, ethnicity, gender) |>
  full_join(kids) |>
  filter(!is.na(included)) |>
  mutate(
    age_days = (date_test - date_birth) |> as.numeric(),
    age_month = floor(age_days / 30.43)
  ) |>
  select(-age_days) |>
  select(expt, id, game, sid, gender, age_month, ethnicity, hispanic) |>
  write_csv(here("data/clean_demog_1_2.csv"))
```