### 2025-10-28 code/expt_3/data_prep_3.Rmd

---
title: "Wrangle transcripts"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)

ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_source <- "data/expt_3/rounds.csv"

player_source <- "data/expt_3/players.csv"

```

# Empirica output prep

```{r}

id_sub <- function(id){
  new_id = case_when(
    id=="pear" ~ "id153",
    id=="apple" ~ "id154",
    id=="pear155" ~ "id155",
    id=="apple156" ~ "id156",
    id=="pear157" ~ "id157",
    id=="apple158" ~ "id158",
    id=="pear123" ~ "id163",
    id=="apple124" ~ "id164",
    id=="pear124"~"id165",
    id=="apple125" ~ "id166",
    T ~ str_c("id", id)
  )
  return(new_id)
}

players <- read_csv(here(player_source)) |> select(playerId=`_id`, id) |> rowwise() |>  mutate(id=id_sub(id)) |> ungroup()

raw_data <- read.csv(here(data_source), header = TRUE) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>%
  pivot_longer(
    cols = starts_with("player"), names_to = c("playerId", "info"), names_prefix = "player_", names_sep = "_",
    values_to = "values", values_transform = as.character, values_drop_na = T
  ) %>%
  filter(values != "") %>%
  filter(playerId != speaker) %>%
  pivot_wider(names_from = info, values_from = values) |>
  left_join(players, by = c("playerId")) |>
  select(-playerId) |>
  rename(playerId = id) |>
  filter(!is.na(playerId)) |>
  left_join(players, by = c("speaker" = "playerId")) |>
  select(-speaker) |>
  rename(speaker = id)

label_image <- function(string) {
  new_string <- case_when(
    string == "/experiment/tangram_E.png" ~ "bunny",
    string == "/experiment/tangram_G.png" ~ "dancer",
    string == "/experiment/tangram_H.png" ~ "baby",
    string == "/experiment/tangram_A.png" ~ "zombie",
    string == "/experiment/banana.png" ~ "banana",
    string == "/experiment/bird.jpeg" ~ "bird",
    string == "/experiment/dog.png" ~ "dog",
    string == "/experiment/apple.png" ~ "apple",
    string == "/experiment/star.png" ~ "star",
    string == "/experiment/crown.png" ~ "crown",
    string == "/experiment/lemon.png" ~ "lemon",
    string == "/experiment/pinwheel.png" ~ "pinwheel"
  )
  return(new_string)
}

less_raw_data <- raw_data %>%
  select(target, distractor, trialNum, targetNum, repNum, response, correct, time, listener = playerId, speaker, gameId) %>%
  mutate(across(c(target, distractor, response), ~ label_image(.x))) |> 
  inner_join(read_tsv(here("data/expt_3/link_transcripts.csv")) |> select(gameId, game)) |> 
  write_csv(here("data/expt_3/clean_data.csv"))

```

# Transcript wrangle

```{r}

wrangle_transcript <- function(game_num) {
  if(file.exists(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))){
  file = read_tsv(here("data/expt_3/raw_transcripts", str_c(game_num, ".tsv")))
  }
  else{
    if(file.exists(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))){
        file = read_tsv(here("data/expt_3/raw_transcripts/exclude", str_c(game_num, ".tsv")))
    }
    else{
      file=tibble(start=NA, end=NA, text=NA, speaker=NA, trial=NA, role=NA, echo=NA, description=NA)
    }
  }
  file|>
    select(start, end, text, speaker, trial, role, echo, description) |>
    fill(trial) %>%
    mutate(start = start / 1000, end = end / 1000) |>
    mutate(trial = trial-1) |>
    nest(data = c(start, end, text, speaker, trial, role, echo, description))

}

transcripts <- read_tsv(here("data/expt_3/link_transcripts.csv")) |> 
  select(gameId, game) |>
  mutate(transcript = map_df(game, wrangle_transcript)) |>
  unnest(transcript) |>
  unnest(data)

write_csv(transcripts, here("data/expt_3/transcripts.csv"))
```

# Sbert data prep

```{r}
transcript <- read_csv(here("data/expt_3/transcripts.csv"))

dat <- read_csv(here("data/expt_3/clean_data.csv")) %>% select(trial = trialNum, gameId, target, distractor)
transcript %>%
  filter(!is.na(description)) %>%
  group_by(trial, gameId) %>%
  mutate(trial = as.numeric(trial)) |>
  filter(trial > 3) |>
  summarize(words = str_c(description, sep = " ", collapse = ", ")) %>%
  left_join(dat) %>%
  write_csv(here("data/expt_3/pre_sbert.csv"))
```

(go run sbert here)
```{r}
library(reticulate)
np <- import("numpy")
mat <- np$load(here("data/expt_3/post_sbert.npy"))
saveRDS(mat, here("data/expt_3/post_sbert.RData"))
```

```{r}
sbert <- read_csv(here("data/expt_3/pre_sbert.csv")) %>% bind_cols(read_rds(here("data/expt_3/post_sbert.RData")) %>%
  as_tibble())

### helper funcs
get_sim_matrix <- function(df, F_mat, method = "cosine") {
  feats <- F_mat[df$feature_ind, ]
  if (method == "cor") {
    return(cor(t(feats), method = "pearson"))
  } else if (method == "euclidean") {
    return(as.matrix(dist(feats, method = "euclidean")))
  } else if (method == "cosine") {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0("unknown method", method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(
      dim1 = as.character(dim1),
      dim2 = as.character(dim2)
    )
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(
      get_sim_matrix(., F_mat, method = method),
      as.character(.$combinedId)
    ))
}

### funcs
F_mat <- sbert %>%
  select(starts_with("V")) %>%
  as.matrix() # Features
M_mat <- sbert %>%
  select(-starts_with("V")) %>%
  mutate(feature_ind = row_number())

similarities <- M_mat %>%
  mutate(combinedId = str_c(gameId, trial, sep = "_")) %>%
  make_across_df(F_mat, "cosine") %>%
  separate(dim1, into = c("game1", "trial1"), convert = T, sep = "_") %>%
  separate(dim2, into = c("game2", "trial2"), convert = T, sep = "_") %>%
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  write_rds(here("data/expt_3/similarities.rds"))
```