### 2025-10-29 code/expt_3/analysis_3.Rmd

---
title: "Kid tg 2"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)
library(brms)
library(rstan)
library(rstanarm)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_loc <- "data/expt_3"
mod_loc <- "code/expt_3/models"
```

# Exclusions

```{r}
transcript <- read_csv(here(data_loc, "transcripts.csv"))
data <- read_csv(here(data_loc, "clean_data.csv"))
link <- read_tsv(here(data_loc, "link_transcripts.csv"))

 exclude_echo <- transcript |>
   filter(!is.na(echo)) |>
   select(game, gameId, trial)

data <- read_csv(here(data_loc, "clean_data.csv")) |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |> 
  anti_join(exclude)

transcript <- transcript |>
    inner_join(link |> filter(is.na(exclude))) |> 
  anti_join(exclude) |>
  filter(!is.na(description))

words <- transcript |> mutate(words=description |> str_count("\\S+")) |>
  filter(role=="S") |>
  group_by(gameId, game, trial) |>
  summarize(words=sum(words))

all_data <- data |>left_join(words) |> 
  filter(!is.na(words)) |> 
   mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |> 
  mutate(correct=as.numeric(correct))

all_data |> select(game, speaker) |> distinct()

```

# Accuracy

```{r}

time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

all_data |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot", geom="point", size=2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = all_data |> filter(trial > 3), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  theme(
    aspect.ratio = 4 / 5,
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```
```{r}
all_data |> filter(type!="practice") |> ggplot(aes(x=reorder(gameId, as.numeric(correct)), y=as.numeric(correct)))+stat_summary(fun.data="mean_cl_boot")+theme(axis.text.x=element_blank())
```

```{r}
all_data |>
  filter(type != "practice") |>
  ggplot(aes(x = type, y = as.numeric(correct), color = target)) +
  theme_bw() +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0, 1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  labs(x = "", y = "Fraction correct") +
  theme(legend.position = "bottom")
```

# Description length

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(all_data, aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = all_data|> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  theme(legend.position = "none")
```

# Speed

```{r}

time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(all_data, aes(x = trial + 1, y = time / 1000, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = all_data |> filter(trial > 3), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  theme(legend.position = "none")
```

# Similarities

```{r}
sims <- read_rds(here(data_loc, "similarities.rds")) |>
  as_tibble() |>
  inner_join(data |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(data |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = trial1 %/% 4, block2 = trial2 %/% 4) |> 
  mutate(same_speaker=(speaker1==speaker2) |> as.numeric(),
         same_game=(game1==game2) |> as.numeric(),
        later = ifelse(block1 > block2, block1, block2), 
        earlier = ifelse(block1 > block2, block2, block1)) |> mutate(target=target1)
```

```{r}
sims |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target1)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  theme(legend.position = "bottom")
```

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```
```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```
```{r}
sims |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

# Run Analyses

```{r}

acc_priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod <- brm(correct ~ trial + (trial | gameId) + (1 | target),
  family = bernoulli(link = "logit"),
  data = all_data |> filter(!is.na(repNum)),
  file = here(mod_loc, "acc"),
  prior = acc_priors,
  control = list(adapt_delta = .95)
)
```

```{r}

red_priors <- c(
  set_prior("normal(5, 10)", class = "Intercept"),
  set_prior("normal(0,5)", class="b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

red_mod <- brm(words ~ trial + (trial | gameId) + (1 | target),
  data = all_data |> filter(!is.na(repNum)),
  file = here(mod_loc, "red"),
  prior = red_priors,
  control = list(adapt_delta = .95)
)
```

```{r}

speed_priors <- c(
  set_prior("normal(60, 100)", class = "Intercept"),
  set_prior("normal(0,20)", class="b"),
  set_prior("normal(0,20)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

speed_mod <- brm(time_seconds ~ trial + (trial | gameId) + (1 | target),
  data = all_data |> filter(!is.na(repNum)) |> mutate(time_seconds=time/1000),
  file = here(mod_loc, "speed"),
  prior = speed_priors,
  control = list(adapt_delta = .95)
)
```

```{r}
sim_priors <- c(
  set_prior("normal(.5, .2)", class = "Intercept"),
  set_prior("normal(0,.1)", class="b"),
  set_prior("normal(0,.05)", class = "sd"))

sim_no_time <- brm(sim ~ same_game + same_speaker + (1|target),
                   data=sims ,
                   file = here(mod_loc, "sim_no_time"),
  prior = sim_priors,
  control = list(adapt_delta = .95))
```

```{r}

sim_to_last <- brm(sim ~ earlier+same_speaker+(1|game)+(1|target),
                   data=sims |> filter(later==4) |> filter(game1==game2) |> mutate(game=game1),
                   file = here(mod_loc, "sim_to_last"),
  prior = sim_priors,
  control = list(adapt_delta = .95))

sim_to_next <- brm(sim ~ earlier+same_speaker+(1|game)+(1|target),
                   data=sims |> filter(later==earlier+1) |> filter(game1==game2) |> mutate(game=game1),
                   file = here(mod_loc, "sim_to_next"),
  prior = sim_priors,
  control = list(adapt_delta = .95))

sim_across <- brm(sim ~ block+(1|target),
                   data=sims |> filter(later==earlier) |> filter(game1!=game2) |> mutate(block=earlier),
                   file = here(mod_loc, "sim_across"),
  prior = sim_priors,
  control = list(adapt_delta = .95))
```

# Analysis results

[x] Logistic model of accuracy: correct ~ trialNum + (trialNum|game) + (1|target) (prior normal(0,1) for beta and sd, lkj(1) for correlation)

```{r}

summary(acc_mod)

```

no effect of trial on accuracy; accuracy generally high

[x] Linear model of description length: words ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(5,10), beta and sd priors of normal(0,5), and correlation of lkj(1))

```{r}

summary(red_mod)

```

no effect of trial on length; generally pretty short throughout

[x] Speed: time_seconds ~ trialNum + (trialNum|game) + (1|target) (prior intercept of normal(60,100), beta and sd priors of (0,20) and correlation of lkj(1))

```{r}

summary(speed_mod)

```

kids get faster over time (by about .7 seconds a pop)

We will interpret the fixed effect of trial number as being about changes in referring over time.

## Sbert models
For utterance similarities, we will embed the describer’s descriptions using S-BERT and use cosine similarity as a proxy for utterance similarity.

We are unsure whether 4 blocks will be sufficient to see potential change over time, so we will

[x]* compare similarities of utterances between different games v within one game v within one speaker: sim ~ same_game + same_speaker + (1|target)

```{r}
summary(sim_no_time)
```

utterances from within the same game are more similar (.34 across games v .59 within), smaller effect of same speaker(+.08, so .67)

[x]* compare similarities of utterances to the last utterance within a game: sim_to_last ~ earlier_block + same_speaker + (1|game) + (1|target)

```{r}
summary(sim_to_last)
```

utterances do get (slightly) more similar to last utterance over time, with a probable bump if same speaker

[x]* compare similarities of utterances to the next utterance: sim_to_next ~ earlier_block + same_speaker + (1|game) + (1|target)

```{r}
summary(sim_to_next)
```

utterances are again more similar from same speaker, but similarity to next is not clearly increasing over time

[x]* compare the similarity level of descriptions from the same block in different games: sim ~ block + (1|target)

```{r}
summary(sim_across)
```

distance of descriptions across games is stable (not seeing divergence)

Priors for the above similarity models: normal(.5,.2) for intercept, normal(0,.1) for
beta, and normal(0,.05) for sd.

We will also qualitatively explore children’s turn-taking structure and how they understand each other.

TODO: the qualitative thing