### 2025-10-29 code/ms_analysis.Rmd

---
title: "Kid tangrams joint analysis"
output:
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F)
library(tidyverse)
library(here)
library(jsonlite)
library(ggplot2)
library(brms)
library(rstan)
library(rstanarm)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep = " "), " ]") %>%
    fromJSON(flatten = T)
}

data_loc_1 <- "data/expt_1"
data_loc_2 <- "data/expt_2"
data_loc_3 <- "data/expt_3"
mod_loc <- "code/mods_for_paper"
```

# Read in data

expt 1

```{r}
transcript <- read_csv(here(data_loc_1, "transcripts.csv"))
data <- read_csv(here(data_loc_1, "clean_data.csv"))
link <- read_csv(here(data_loc_1, "link_transcripts.csv"))

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude)))

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_1 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 2 ~ "practice",
    trial < 6 ~ "block 1",
    trial < 10 ~ "block 2",
    trial < 14 ~ "block 3",
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "1")
```

expt 2

```{r}
transcript <- read_csv(here(data_loc_2, "transcripts.csv"))
data <- read_csv(here(data_loc_2, "clean_data.csv"))
link <- read_csv(here(data_loc_2, "link_transcripts.csv"))

exclude_echo <- transcript |>
  filter(!is.na(echo)) |>
  select(game, gameId, trial)

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo)

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_2 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "2")
```

expt 3

```{r}
transcript <- read_csv(here(data_loc_3, "transcripts.csv"))
data <- read_csv(here(data_loc_3, "clean_data.csv"))
link <- read_tsv(here(data_loc_3, "link_transcripts.csv")) |> select(gameId, game, exclude, exclusion_reason)

exclude_echo <- transcript |>
  filter(!is.na(echo)) |>
  select(game, gameId, trial)

data <- data |>
  rename(trial = trialNum) |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo)

transcript <- transcript |>
  inner_join(link |> filter(is.na(exclude))) |>
  anti_join(exclude_echo) |>
  filter(!is.na(description))

words <- transcript |>
  mutate(words = description |> str_count("\\S+")) |>
  filter(role == "S") |>
  group_by(gameId, game, trial) |>
  summarize(words = sum(words))

all_words <- transcript |> 
  filter(on_topic=="Y") |> 
  mutate(words = text |> str_count("\\S+")) |>
  group_by(gameId, game, trial) |>
  summarize(all_words = sum(words))

all_data_3 <- data |>
  left_join(words) |>
  left_join(all_words) |> 
  filter(!is.na(words)) |>
  mutate(type = case_when(
    trial < 4 ~ "practice",
    trial < 8 ~ "block 1",
    trial < 12 ~ "block 2",
    trial < 16 ~ "block 3",
    trial < 20 ~ "block 4"
  )) |>
  filter(!is.na(response)) |>
  mutate(correct = as.numeric(correct)) |>
  mutate(expt = "3")
```

```{r}
demogs <- read_csv(here(data_loc_1, "clean_demog.csv")) |>
  bind_rows(read_csv(here(data_loc_2, "clean_demog.csv"))) |>
  bind_rows(read_csv(here(data_loc_3, "clean_demog.csv"))) |>
  mutate(expt = as.character(expt))
```

```{r}
joint <- all_data_1 |>
  bind_rows(all_data_2) |>
  bind_rows(all_data_3)

speaker_demogs <- demogs |> select(speaker = id, expt, game, speaker_gender = gender, speaker_age = age_month)

listener_demogs <- demogs |> select(listener = id, expt, game, listener_gender = gender, listener_age = age_month)

joint_demog <- joint |>
  left_join(speaker_demogs) |>
  left_join(listener_demogs) |>
  mutate(trial = case_when(
    expt=="1"~ trial-2,
    T ~ trial-4,
  ),     time.sec=time/1000
) |> filter(type!="practice") |> select(-time) |> 
  write_csv(here("data/processed_for_paper/joint_with_demog.csv"))
```

```{r}
sims_1 <- read_rds(here(data_loc_1, "similarities.rds")) |>
  as_tibble() |> 
  inner_join(all_data_1 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_1 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |> 
  filter(target1 == target2) |> 
  mutate(block1 = (trial1-2) %/% 4, block2 = (trial2-2) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "1")

sims_2 <- read_rds(here(data_loc_2, "similarities.rds")) |>
  as_tibble() |>
  inner_join(all_data_2 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_2 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = (trial1-4) %/% 4, block2 = (trial2-4) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "2")

sims_3 <- read_rds(here(data_loc_3, "similarities.rds")) |>
  as_tibble() |>
  inner_join(all_data_3 |> select(game1 = gameId, trial1 = trial, target1 = target, speaker1 = speaker)) |>
  inner_join(all_data_3 |> select(game2 = gameId, trial2 = trial, target2 = target, speaker2 = speaker)) |>
  filter(target1 == target2) |>
  mutate(block1 = (trial1-4) %/% 4, block2 = (trial2-4) %/% 4) |>
  mutate(
    same_speaker = (speaker1 == speaker2) |> as.numeric(),
    same_game = (game1 == game2) |> as.numeric(),
    later = ifelse(block1 > block2, block1, block2),
    earlier = ifelse(block1 > block2, block2, block1)
  ) |>
  mutate(target = target1) |>
  mutate(expt = "3")

sims <- sims_1 |>
  bind_rows(sims_2) |>
  bind_rows(sims_3) |>
  select(-target1, -target2) |> 
  write_csv(here("data/processed_for_paper/sims.csv"))
```

# QC

TODO TURN INTO ASSERTIONS
```{r, eval=F}
joint_demog |>
  select(speaker, game, expt) |>
  distinct() |>
  group_by(game, expt) |>
  tally() |>
  filter(n != 2)

joint_demog |>
  select(listener, game, expt) |>
  distinct() |>
  group_by(game, expt) |>
  tally() |>
  filter(n != 2)

joint_demog |>
  select(speaker, game, expt, trial) |>
  distinct() |>
  group_by(game, expt, trial) |>
  tally() |>
  filter(n != 1)

joint_demog |>
  select(listener, game, expt, trial) |>
  distinct() |>
  group_by(game, expt, trial) |>
  tally() |>
  filter(n != 1)
# targetNum
# repNum
```

```{r, eval=F}
transcript_1 <- read_csv(here(data_loc_1, "transcripts.csv")) |>
  left_join(read_csv(here(data_loc_1, "link_transcripts.csv"))) |>
  mutate(expt = "1")

transcript_2 <- read_csv(here(data_loc_2, "transcripts.csv")) |>
  left_join(read_csv(here(data_loc_2, "link_transcripts.csv"))) |>
  mutate(expt = "2")

transcript_3 <- read_csv(here(data_loc_3, "transcripts.csv")) |>
  left_join(read_tsv(here(data_loc_3, "link_transcripts.csv"))) |>
  mutate(expt = "3")

transcripts <- transcript_1 |>
  bind_rows(transcript_2) |>
  bind_rows(transcript_3)

# confirm that all S and L tags go with kids

transcripts |>
  filter(role %in% c("S", "L")) |>
  filter(!str_detect(speaker, "id")) |>
  select(expt, game, speaker, role, text)

transcripts |>
  select(trial, expt) |>
  distinct() # should be 0-19 (0-13 for expt 1) + NAs

transcripts |>
  filter(!is.na(trial)) |>
 filter(str_detect(speaker, "id")) |>
  filter(is.na(role)) |>
  select(expt, game, trial, text) |> View() # to check on later!!!

transcripts |> filter(!is.na(trial)) |> 
    filter(!is.na(role)) |>
  select(game, expt, trial, speaker, role) |> 
  distinct()|>group_by(game, expt, trial, speaker) |> tally() |> filter(n!=1)

transcripts |> filter(!is.na(trial)) |> 
    filter(!is.na(role)) |>
  select(game, expt, trial, speaker, role) |> 
  distinct()|>group_by(game, expt, trial, role) |> tally() |> filter(n!=1)

transcripts |>
  filter(!is.na(trial)) |>
  filter(!is.na(role)) |>
  select(game, expt, trial, speaker_transcript = speaker, role, text) |>
  filter(role == "S") |>
  full_join(joint_demog) |>
  filter(speaker != speaker_transcript) |>
  select(expt, game, trial, text, speaker_transcript, speaker, listener)
```

# Graphs
## Accuracy
```{r, fig.width=5, fig.height=5}
time_palette <- c("practice" = "grey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

joint_demog |>
  ggplot(aes(x = trial + 1, y = correct)) +
  stat_summary(aes(color = type), fun.data = "mean_cl_boot", geom = "point", size = 2) +
  stat_summary(fun.data = "mean_cl_boot", geom = "linerange", color = "black") +
  scale_color_manual(values = time_palette) +
  # geom_bar(position = position_dodge(.9), color = "black", fill = "#19ADDE", width = .7) +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  # geom_text(aes(label = percent), hjust = 1.6, color = "white", size = 3) +
  theme_classic() +
  xlab("\nTrial") +
  ylab("Percent accuracy\n") +
  scale_x_continuous(breaks = seq(1, 20, 1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  geom_hline(yintercept = .5) +
  facet_grid(expt ~ .) +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 9, color = "black"),
    legend.position = "none"
  )
```

```{r}
joint_demog |>
  filter(type != "practice") |>
  ggplot(aes(x = type, y = as.numeric(correct), color = target)) +
  theme_bw() +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1), limits = c(0, 1), expand = c(0, 0)) +
  geom_hline(yintercept = .5) +
  labs(x = "", y = "Fraction correct") +
  theme(legend.position = "bottom")
```

## Speed

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = time.sec, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Time (seconds)") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

## Reduction

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

```{r}
time_palette <- c("practice" = "darkgrey", "block 1" = "#E41A1C", "block 2" = "#377EB8", "block 3" = "#4DAF4A", "block 4" = "purple") # TODO FIX

ggplot(joint_demog, aes(x = trial + 1, y = all_words, color = type)) +
  geom_jitter(alpha = .5, color = "grey") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = seq(1, 20, 1), expand = c(0, 0)) +
  scale_color_manual(values = time_palette) +
  theme_classic() +
  geom_smooth(data = joint_demog |> filter(type != "practice"), method = "lm", color = "black") +
  xlab("Trial") +
  ylab("Words") +
  facet_grid(expt ~ .) +
  theme(legend.position = "none")
```

# Models
## Accuracy
```{r}
acc_priors <- c(
  set_prior("normal(0,1)", class = "b"),
  set_prior("normal(0,1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod_1 <- brm(correct ~ trial + (trial | game) + (1 | target),
                 family = bernoulli(link = "logit"),
                 data = joint_demog |> filter(expt=="1"),
                 file = here(mod_loc, "acc_1.rds"),
                 prior = acc_priors,
                 control = list(adapt_delta = .95)
)

acc_mod_2 <- brm(correct ~ trial+ (trial| gameId) + (1 | target),
               family = bernoulli(link = "logit"),
               data = joint_demog |> filter(expt=="2"),
               file = here(mod_loc, "acc_2.rds"),
               prior = acc_priors,
               control = list(adapt_delta = .95)
)

acc_mod_3 <- brm(correct ~ trial + (trial | gameId) + (1 | target),
               family = bernoulli(link = "logit"),
               data = joint_demog |> filter(expt=="3"),
               file = here(mod_loc, "acc_3.rds"),
               prior = acc_priors,
               control = list(adapt_delta = .95)
)

acc_mod_all <- brm(correct ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                    family = bernoulli(link = "logit"),
                    data = joint_demog,
                    file = here(mod_loc, "acc_meta.rds"),
                    prior = acc_priors,
                    control = list(adapt_delta = .95)
)

```

## Speed
```{r}
speed_priors <- c(
  set_prior("normal(60,100)", class = "Intercept"),
  set_prior("normal(0,20)", class = "b"),
  set_prior("normal(0,20)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

speed_mod_1 <- brm(time.sec ~ trial + (trial | game) + (1 | target),
                   data = joint_demog |> filter(expt=="1"),
                   file = here(mod_loc, "speed_1.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_2 <- brm(time.sec ~ trial + (trial | gameId) + (1 | target),
                   data = joint_demog |> filter(expt=="2"),
                   file = here(mod_loc, "speed_2.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_3 <- brm(time.sec ~ trial + (trial | gameId) + (1 | target),
                   data = joint_demog |> filter(expt=="3"),
                   file = here(mod_loc, "speed_3.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)

speed_mod_meta <- brm(time.sec ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                   data = joint_demog,
                   file = here(mod_loc, "speed_meta.rds"),
                   prior = speed_priors,
                   control = list(adapt_delta = .95)
)
```

## Description length

```{r}
description_priors <- c(
  set_prior("normal(5,10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0,5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

description_mod_1 <- brm(words ~ trial + (trial | game) + (1 | target),
                         data = joint_demog |> filter(expt=="1"),
                         file = here(mod_loc, "description_1.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_2 <- brm(words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="2"),
                         file = here(mod_loc, "description_2.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_3 <- brm(words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="3"),
                         file = here(mod_loc, "description_3.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

description_mod_meta <- brm(words ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                            data = joint_demog,
                            file = here(mod_loc, "description_meta.rds"),
                            prior = description_priors,
                            control = list(adapt_delta = .95)
)
```

## Description -- robustness
we can instead of just the cleaned up description, look at all kids on topic words (including listener ?s, self repeitions, saying I don't know...)

this excludes off topic things (I have a stuffy too / comments about smurfy), stuff about the mechanics of the game (do I press it?), and celebration/non-referential commentary on how the game is going (yay!, I'm winning. that was wrong)

TODO 
```{r}
description_priors <- c(
  set_prior("normal(5,10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0,5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

all_description_mod_1 <- brm(all_words ~ trial + (trial | game) + (1 | target),
                         data = joint_demog |> filter(expt=="1"),
                         file = here(mod_loc, "all_description_1.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_2 <- brm(all_words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="2"),
                         file = here(mod_loc, "all_description_2.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_3 <- brm(all_words ~ trial + (trial | gameId) + (1 | target),
                         data = joint_demog |> filter(expt=="3"),
                         file = here(mod_loc, "all_description_3.rds"),
                         prior = description_priors,
                         control = list(adapt_delta = .95)
)

all_description_mod_meta <- brm(all_words ~ trial + (trial | gameId) + (1 | target) + (1|expt),
                            data = joint_demog,
                            file = here(mod_loc, "all_description_meta.rds"),
                            prior = description_priors,
                            control = list(adapt_delta = .95)
)
```

# S-BERT graphs

## Sim to self/partner/others
```{r}
sims |>
  mutate(source = case_when(
    speaker1 == speaker2 ~ "same speaker",
    game1 == game2 ~ "same game",
    T ~ "different games"
  )) |>
  ggplot(aes(x = source, y = sim, color = target)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(y = "Cosine similarity", x = "Comparing two utterances from ...", color = "target image") +
  facet_wrap(~expt) +
  theme(legend.position = "bottom")
```

## Sim to last round

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_grid(expt ~ same_speaker) +
  labs(x = "Earlier block to last block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## Sim to next round

```{r}
sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "same speaker", "different speaker")) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## Sim across game

```{r}
sims |>
  filter(game1 != game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier) |>
  ggplot(aes(as.character(earlier), sim, color = target)) +
  geom_point(alpha = .01, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Cross game similarity in block ...", y = "cosine similarity", color = "target") +
  theme(legend.position = "bottom")
```

## similarity and accuracy

```{r}
sim_acc_to_next <- sims |> 
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
    earlier_block.num = earlier
  ) |> 
  select(earlier_trial, target, game1, earlier, later, sim, earlier_block.num, same_speaker, expt) |> left_join(joint_demog |> select(target, earlier_trial=trial, game1=gameId, correct, expt)) |> filter(!is.na(correct))

sim_acc_to_next |> ggplot(aes(as.character(earlier), sim, color = as.character(correct))) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  labs(x = "Earlier block to *next* block", y = "cosine similarity") +
  theme(legend.position = "bottom")+facet_wrap(~expt)
```

# S-BERT models

# across-between 
```{r}
conv_priors <- c(
  set_prior("normal(.5,.2)", class = "Intercept"),
  set_prior("normal(0,.1)", class = "b"),
  set_prior("normal(0,.05)", class = "sd")
)

sim_across_between_data <- sims |>
  mutate(same_game = case_when(
    game1 == game2 ~ 1,
    T ~ 0
  )) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, 1, 0))

sim_across_between_mod_1 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="1"),
                                file = here(mod_loc, "sim_across_between_1.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_2 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="2"),
                                file = here(mod_loc, "sim_across_between_2.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_3 <- brm(sim ~ same_game + same_speaker + (1 | target),
                                data = sim_across_between_data |> filter(expt=="3"),
                                file = here(mod_loc, "sim_across_between_3.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

sim_across_between_mod_meta <- brm(sim ~ same_game + same_speaker + (1 | target) + (1|expt),
                                   data = sim_across_between_data,
                                   file = here(mod_loc, "sim_across_between_meta.rds"),
                                   prior = conv_priors,
                                   control = list(adapt_delta = .95)
)
```

## to last

```{r}
sim_to_last_data <- sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 3) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
  )

sim_to_last_mod_2 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_last_data |> filter(expt=="2"),
                         file = here(mod_loc, "sim_to_last_2.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
sim_to_last_mod_3 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_last_data |> filter(expt=="3"),
                         file = here(mod_loc, "sim_to_last_3.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_last_mod_meta <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target)+ (1|expt),
                         data = sim_to_last_data,
                         file = here(mod_loc, "sim_to_last_meta.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
```

## to next

```{r}
sim_to_next_data <- sims |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0),
  )

sim_to_next_mod_2 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_next_data |> filter(expt=="2"),
                         file = here(mod_loc, "sim_to_next_2.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_next_mod_3 <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target),
                         data = sim_to_next_data |> filter(expt=="3"),
                         file = here(mod_loc, "sim_to_next_3.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)

sim_to_next_mod_meta <- brm(sim ~ earlier + same_speaker + (1 | game1) + (1 | target)+ (1|expt),
                         data = sim_to_next_data,
                         file = here(mod_loc, "sim_to_next_meta.rds"),
                         prior = conv_priors,
                         control = list(adapt_delta = .95)
)
```

## cross game

```{r}
sim_across_data <- sims |>
  filter(game1 != game2) |>
  filter(block1 == block2)

sim_across_mod_2 <- brm(sim ~ block1 + (1 | target),
                        data = sim_across_data |> filter(expt=="2"),
                        file = here(mod_loc, "sim_across_2.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

sim_across_mod_3 <- brm(sim ~ block1 + (1 | target),
                        data = sim_across_data |> filter(expt=="3"),
                        file = here(mod_loc, "sim_across_3.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

sim_across_mod_meta <- brm(sim ~ block1 + (1 | target)+ (1|expt),
                        data = sim_across_data,
                        file = here(mod_loc, "sim_across_meta.rds"),
                        prior = conv_priors,
                        control = list(adapt_delta = .95)
)

```

## similarity & accuracy

may need to double check this one!
```{r}

sim_acc_to_next <- sims |> 
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  mutate(earlier_trial = ifelse(trial1 > trial2, trial2, trial1)) |>
  filter(later == earlier + 1) |>
  mutate(
    same_speaker = ifelse(speaker1 == speaker2, 1, 0)
  ) |> 
  select(target, game1, earlier, later, sim, same_speaker, expt) |> 
left_join(joint_demog |> select(target, earlier=repNum, game1=gameId, correct, expt ))

conv_priors <- c(
  set_prior("normal(.5,.2)", class = "Intercept"),
  set_prior("normal(0,.1)", class = "b"),
  set_prior("normal(0,.05)", class = "sd")
)

sim_acc_to_next_mod_meta <- brm(sim ~ earlier*correct + same_speaker*correct + (1 | game1) + (1 | target) + (1|expt),
                                data = sim_acc_to_next,
                                file = here(mod_loc, "sim_acc_to_next_meta.rds"),
                                prior = conv_priors,
                                control = list(adapt_delta = .95)
)

```

# Demog graphs

## Accuracy
```{r}
joint_demog |> ggplot(aes(x = speaker_age, y = correct, color = interaction(speaker_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(correct=mean(correct, na.rm=T)))

joint_demog |> ggplot(aes(x = listener_age, y = correct, color = interaction(listener_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(listener_age, expt, game, listener_gender) |> summarize(correct=mean(correct, na.rm=T)))

joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = correct, color = as.character(girlness))) +
  labs(x="Mean age of participants", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(correct=mean(correct, na.rm=T))) +
  geom_smooth(method = "lm")
```

## words

```{r}
joint_demog |> ggplot(aes(x = speaker_age, y = words, color = interaction(speaker_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(speaker_age, speaker, expt, game, speaker_gender) |> summarize(words=mean(words, na.rm=T)))

joint_demog |> ggplot(aes(x = listener_age, y = words, color = interaction(listener_gender))) +
  geom_smooth(method = "lm")+
  geom_jitter(data=joint_demog |> group_by(listener_age, expt, game, listener_gender) |> summarize(words=mean(words, na.rm=T)))

joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |>
  ggplot(aes(x = mean_age, y = words, color = as.character(girlness))) +
  labs(x="Mean age of participants", y="Accuracy", color="Number of girls")+
  geom_jitter(data=joint_demog |>
  mutate(mean_age = (listener_age + speaker_age) / 2, girlness = (listener_gender == "female") + (speaker_gender == "female")) |> group_by(expt, game, mean_age, girlness) |> summarize(words=mean(words, na.rm=T))) +
  geom_smooth(method = "lm")
```

# Demog models
## Accuracy

```{r}

mean_age <- joint_demog |> distinct(expt, speaker, speaker_age) |> summarize(mean_age=mean(speaker_age)) |> pluck(1,1)

centered_demog <- joint_demog |> 
  mutate(
    centered_speaker_age = speaker_age - mean_age,
    centered_listener_age = listener_age - mean_age,
    speaker_gender_num = ifelse(speaker_gender == "female", .5, -.5),
    listener_gender_num = ifelse(listener_gender == "female", .5, -.5)
  )

acc_priors <- c(
  set_prior("normal(0, 1)", class = "b"),
  set_prior("normal(0, 1)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

acc_mod_demog <- brm(
  correct ~ trial +
    centered_speaker_age + speaker_gender_num +
    centered_listener_age + listener_gender_num +
    (trial | game) + (1 | target),
  family = bernoulli(link = "logit"),
  data = centered_demog,
  file = here(mod_loc, "acc_demog.rds"),
  prior = acc_priors,
  control = list(adapt_delta = .95)
)

```

## Description length

```{r}
red_priors <- c(
  set_prior("normal(5, 10)", class = "Intercept"),
  set_prior("normal(0,5)", class = "b"),
  set_prior("normal(0, 5)", class = "sd"),
  set_prior("lkj(1)", class = "cor")
)

red_mod <- brm(
  words ~ trial*(centered_speaker_age + speaker_gender_num +
    centered_listener_age + listener_gender_num) +
    (trial | game) + (1 | target),
  data = centered_demog,
  file = here(mod_loc, "description_demog.rds"),
  prior = red_priors,
  control = list(adapt_delta = .95)
)

```

# Save model summaries

```{r}
library(tidybayes)

save_summary <- function(model) {
  intervals <- gather_draws(model, `b_.*`, regex = T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex = T) %>%
    mutate(above_0 = ifelse(.value > 0, 1, 0)) %>%
    group_by(.variable) %>%
    summarize(pct_above_0 = mean(above_0)) %>%
    left_join(intervals, by = ".variable") %>%
    mutate(
      lower = .lower,
      upper = .upper,
      Term = str_sub(.variable, 3, -1),
      Estimate = .value
    ) %>%
    select(Term, Estimate, lower, upper)

  stats
}

do_model <- function(path) {
  model <- read_rds(here(mod_loc, path))
  save_summary(model) |> write_rds(here(mod_loc, "summary", path))
  model$formula |> write_rds(here(mod_loc, "formulae", path))
  print(summary(model))
}

mods <- list.files(path = here(mod_loc), pattern = ".*rds") |> walk(~ do_model(.))
```